---
title: "Visualization"
author: "Christina Chen"
date: "2026-10-01"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

Now we have everything we need!

```{r}
weather_df
```

Let's make a scatterplot!
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```


```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

I always like the dataframe first.

```{r}
ggp_temp_scattterplot = 
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Let's make this a bit fancier...
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) +
  geom_smooth(se = FALSE)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name),alpha = .25) +
  geom_smooth(se = FALSE)
```

Show faceting 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) +
  facet_grid(. ~ name)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) +
  facet_grid(cols = vars(name))
```

Let's look at something else. 

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

make a plot of central park tmax v tmin only, and convert temperatures to fahrenheit.

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax = tmax * (9 / 5) + 32,
    tmin = tmin * (9 / 5) + 32
  ) |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

what's a hex plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")

weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Density plots are great!!

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

Boxplots

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin plots (looks a bit like combined boxplot and density plot)
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

Ridge plots ...

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density()

weather_df |> 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## Save some of my plots

```{r}
ggp_weather =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + 
  facet_grid(. ~ name)

ggp_weather

ggsave("images/ggp_weather.pdf", ggp_weather)
```

```{r fig.width = 12, fig.height = 12}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

## Remember this plot ..?
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .05)
```

## Labels
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .05) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  )
```















