---
title: "Visualization"
author: "Christina Chen"
date: "2026-10-01"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)
data("weather_df")
```

Now we have everything we need!

```{r}
weather_df
```

Let's make a scatterplot!
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```


```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

I always like the dataframe first.

```{r}
ggp_temp_scattterplot = 
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Let's make this a bit fancier...
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) +
  geom_smooth(se = FALSE)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name),alpha = .25) +
  geom_smooth(se = FALSE)
```

Show faceting 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) +
  facet_grid(. ~ name)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .25) +
  facet_grid(cols = vars(name))
```

Let's look at something else. 

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

make a plot of central park tmax v tmin only, and convert temperatures to fahrenheit.

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax = tmax * (9 / 5) + 32,
    tmin = tmin * (9 / 5) + 32
  ) |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

what's a hex plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")

weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Density plots are great!!

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

Boxplots

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin plots (looks a bit like combined boxplot and density plot)
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

Ridge plots ...

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density()

weather_df |> 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## Save some of my plots

```{r}
ggp_weather =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + 
  facet_grid(. ~ name)

ggp_weather

ggsave("images/ggp_weather.pdf", ggp_weather)
```

```{r fig.width = 12, fig.height = 12}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

## Remember this plot ..?
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

## Labels
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  )
```

## Scales

Start with the same plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

Look at color scales

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 300))
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```


## Themes

Shift the legend.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

Change the overall theme.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.esp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in 'geom'

```{r}
central_park =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(data = molokai, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

## `patchwork`

remember faceting?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) +
facet_grid(. ~ name)
```

What happens when you want multipanel plots but can't facet ...?

```{r}
tmax_tmin_p =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p =
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5) +
  theme(legend.position = "none")

tmax_date_p =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p

```

## Data manipulation

Control your factors.

```{r}
weather_df |> 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Molokai_HI"))
  ) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df |> 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) |> 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

## Revisit the pups

Data from the FAS study.

```{r}
pup_data =
  read_csv("data/FAS_pups.csv", skip = 3) |> 
  janitor::clean_names() |> 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_data =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litters_data, by = "litter_number")

fas_data |> 
  select(dose, day_of_tx, starts_with("pd_")) |>
  mutate(pd_ears = as.numeric(pd_ears)) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) |> 
  mutate(outcome = forcats::fct_relevel(outcome, "pd_ears", "pd_pivot", "pd_walk", "pd_eyes")) |> 
  drop_na() |> 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)
```




























