---
title: "Exploratory analysis"
author: "Christina Chen"
date: "2026-10-08"
output: github_document
---
```{r message = F}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Load the weather data and round date to the first day of the month

```{r}
data("weather_df")

weather_df =
  weather_df |> 
  mutate(month = lubridate::floor_date(date, unit = "month"))
```

## `group_by`

```{r}
weather_df |> 
  group_by(name, month) |> 
  ungroup(month)
```

## counting things

count month / name observations

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(n_obs = n())
```

we can use `count()`

```{r}
weather_df |> 
  count(name, month, name = "n_obs")
```

**NEVER** use base R's `table`

```{r eval = F}
weather_df |> 
  pull(month) |> 
  table()
```
* The result of this ends up not being a dataframe.
* To get to this table, you have to pull values out, which we do not like doing.


other helpful counters

```{r}
weather_df |> 
  group_by(month) |> 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

## A digression on 2x2 tables

```{r}
weather_df |> 
  filter(name != "Molokai_HI") |> 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) |> 
  group_by(name, cold) |> 
  summarize(count = n())
```

```{r}
weather_df |> 
  filter(name != "Molokai_HI") |> 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) |> 
  janitor::tabyl(name, cold)
```

## General summaries

You can do lots of summaries.

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe! So we can make a plot.

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Suppose you want to summarize many columns.

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(across(prcp:tmin, mean))
```

Reminder: sometimes your results are easier to read in another format.

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(mean_tmax = mean(tmax)) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |> 
  knitr::kable(digits = 1)
```









